Online {\IterComp}

### 2017-09-16 infra.tex
final

Although {\itercomp} had been originally proposed as an \textit{offline} optimization strategy, it can be adapted to work in
\textit{online} scenarios. Instead of selecting the best optimization sequence at development time, a first version of the program is
shipped bundled with the {\itercomp} mechanism. Our mechanism produces new application binaries using different optimization sequences.
When the user starts the application, we load and execute one of these binaries to evaluate the corresponding sequence online. Our
mechanism operates on the application's Intermediate Representation (IR) produced by LLVM. This saves us the effort of repeatedly parsing
the code through the frontend and removes the need for shipping the full application code.

>>>>

Although {\itercomp} had been originally proposed as an \textit{offline} optimization strategy, this work shows that this powerful
technique can be adapted to work in \textit{online} scenarios. Instead of selecting the best optimization sequence at development time, a
first version of the program is shipped bundled with the {\itercomp} mechanism. Our mechanism produces new application binaries using
different optimization sequences. When the user starts the application, we load and execute one of these binaries to evaluate the
corresponding sequence online. Our mechanism operates on the application's Intermediate Representation (IR) produced by LLVM. This saves us
the effort of repeatedly parsing the code through the frontend and removes the need for shipping the full application code.

### 2017-09-16 prof.tex
10 pages
minor
final

\paragraph{Populating the edge flows:}
Intuitively, if all the edge flows are known for the complement of a spanning tree then at any leaf of the spanning tree there is only one
unknown edge flow. This unknown edge flow can be calculated by Kirchhoff's first law\cite{knuth73,ball94},
also known as the principle of \textit{conservation of flow}, which states that
the amount of flow into a vertex equals the amount of output flow.
This process repeats as a bottom-up propagation until all the unknown edge flows have been calculated. This algorithm can be formally defined as a post-order traversal on the
spanning tree. Let $G$ be the CFG. This algorithm first initializes the edge flows $D_{(u,v)}$, for each edge $(u,v)$ in the CFG, as
follows:
\[
D_{(u,v)} \gets
\begin{cases}
    P_{(u,v)} & \quad \text{if edge $(u,v)$ has a probe $P_{(u,v)}$}\\
    0       & \quad \text{otherwise}
\end{cases}
\]
This initialization represents only the trivially known edge flows, which are the
probes themselves (the highlighted edges in red).

@@

\paragraph{Composing the aggregated value of the probes:}
If $x$ is a symbolic expression, then $\kappa_P(x)$ represents the coefficient
of the term $P$ in $x$.
In our case, $\kappa_P(x)$ will belong to the set $\{-1,0,1\}$.
Once all the edge flows are known, then, 
for all probes $P$, the amount of work associated
with probe $P$ is:
\[
%\omega(P) = \sum_{u\in G} \kappa_P(\Sigma^+_u)w(u)
\omega(P) = \sum_{u\in G} \kappa_P(\mathrm{SumOut}(u))w(u)
\]
It would be equally valid to define $\omega(P)$ in terms of $\mathrm{SumIn}(u)$
instead of $\mathrm{SumOut}(u)$,
since these values are equal after populating all edge flows,
as it is enforced by the Kirchhoff's law.

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.8]{figs/cfg-relax-example.pdf}
  \caption{Example of a CFG containing a loop and its decomposition into DAGs.
           The DAGs are the subgraphs within the dashed boundaries.}
  \label{fig:cfg-relax-example}
\end{figure}

>>>>

\paragraph{Populating the edge flows.}
Intuitively, if all the edge flows are known for the complement of a spanning tree, then at any leaf of the spanning tree there is only one
unknown edge flow. This unknown edge flow can be calculated by Kirchhoff's first law~\cite{knuth73,ball94}, which is also known as the
principle of \textit{conservation of flow}. Specifically, this theory states that the amount of flow into a vertex equals the amount of
output flow. To populate the edge flows, we repeat this process as a bottom-up propagation until all the unknown edge flows have been
calculated. This algorithm can be formally defined as a post-order traversal on the spanning tree. Let $G$ be the CFG. This algorithm first
initializes the edge flows $D_{(u,v)}$, for each edge $(u,v)$ in the CFG, as follows:
\[
D_{(u,v)} \gets
\begin{cases}
    P_{(u,v)} & \quad \text{if edge $(u,v)$ has a probe $P_{(u,v)}$}\\
    0       & \quad \text{otherwise}
\end{cases}
\]
This initialization represents only the trivially known edge flows, which are the probes themselves, e.g. the highlighted edges in
Figure~\ref{fig:cfg-example}.

@@

\paragraph{Composing the aggregated value of the probes.}
If $x$ is a symbolic expression, then $\kappa_P(x)$ represents the coefficient
of the term $P$ in $x$.
In our case, $\kappa_P(x)$ will belong to the set $\{-1,0,1\}$.
Once all the edge flows are known, then,
for all probes $P$, the amount of work associated
with probe $P$ is:
\[
%\omega(P) = \sum_{u\in G} \kappa_P(\Sigma^+_u)w(u)
\omega(P) = \sum_{u\in G} \kappa_P(\mathrm{SumOut}(u))w(u)
\]
It would be equally valid to define $\omega(P)$ in terms of $\mathrm{SumIn}(u)$
instead of $\mathrm{SumOut}(u)$,
since these values are equal after populating all edge flows,
as it is enforced by the Kirchhoff's law.

@@

\begin{figure}[t]
  \centering
  \includegraphics[scale=0.8]{figs/cfg-relax-example.pdf}
  \caption{Example of a CFG containing a loop and its decomposition into DAGs.
           The DAGs are the subgraphs within the dashed boundaries.}
  \vspace{-4mm}
  \label{fig:cfg-relax-example}
\end{figure}

### 2017-09-16 related.tex
final

\subsection{{\IterComp}}
Various adaptive compilation techniques have been proposed for finding near optimal compilation
settings~\cite{agakov06,kulkarni04,stephenson03,hoste08}. These approaches compare the performance of candidate compiler options by first
compile the program with different options and then profiling the different compiled versions on the same program input. All these
approaches implicitly assume that the best-performing compiler settings found from a small set of inputs will work well on unseen datasets.
Chen~\etal~\cite{chen10,chen12a} show that the best optimization options can differ across inputs. Their results suggest that, in most
cases it is possible to find a single set of optimizations that produces good results across all inputs; but achieving this requires
developers provide a few tens of representative inputs to be used by the iterative compilation search process. We argue that providing
representative inputs is difficult in practice, as this require developers to foresee all possible user inputs. Even this is possible,
current approaches all require to evaluate each candidate compiler option on every single input. Doing so incurs significant overhead,
which thus prevents iterative compilation to be adopted at a larger scale.

@@

Recent studies~\cite{chen12b,fang15} have used real data to perform {\itercomp} on distributed data center applications. Each worker
receives a subset of the input dataset to evaluate a small set of optimization settings on. The best such setting from each round is used
for subsequent executions of the same code. By testing new settings and re-evaluating old ones on new datasets this approach makes it more
likely to select an optimization setting that works well across inputs. Like in~\cite{chen10} every setting is tested on the same set of
inputs, wasting time and energy. This approach only works well with MapReduce-like workloads, since it relies on the MapReduce framework
for repeating the same computation multiple times without side-effects or corrupting the input.

Fursin~\etal~\cite{fursin07} attempted to find a general approach for online \itercomp. They proposed to use instructions per cycle (IPC)
as a metric for comparing different runs of the same program with different optimization settings and different inputs. Their hypothesis is
that application binaries displaying higher IPC are more efficient than lower IPC binaries. However, many optimization techniques affect
IPC in the opposite way, making the program more efficient while lowering IPC. We have also shown in Section~\ref{sec:motivation}, IPC is
not strongly correlate to speedup. Therefore, it is not a good metric for \itercomp. Furthermore, IPC cannot be used to estimate
performance for multi-threaded applications~\cite{alameldeen06,eyerman08} due to the side-effect of thread contention. Aalameldeen and
Wood~\cite{alameldeen06} also suggest that relying on IPC can be harmful, and work metrics like time per work unit or work units per time
should be used. However, prior works all rely on developers to provide such metrics. This paper presents the first automatic approach to
derive such work metrics from arbitrary code without developer intervention.

>>>>

\subsection{{\IterComp}}
Various adaptive compilation techniques have been proposed for finding near optimal compilation
settings~\cite{agakov06,kulkarni04,stephenson03,hoste08}. These approaches compare the performance of candidate compiler options by first
compile the program with different options and then profiling the different compiled versions on the same program input. All these
approaches implicitly assume that the best-performing compiler settings found from a small set of inputs will work well on unseen datasets.
Chen~\etal~\cite{chen10,chen12a} show that the best optimization options, however, can differ across inputs. Their results suggest that, in
most cases it is possible to find a single set of optimizations that produces good results across all inputs; but achieving this requires
developers provide a few tens of representative inputs to be used by the iterative compilation search process. We argue that providing
representative inputs is difficult in practice, as users may not use the applications as the developer expects. Even this is possible,
current approaches all require to evaluate all candidate compiler options on the same input. Doing so across a range of inputs would incur
significant overhead, which thus prevents iterative compilation to be adopted at a larger scale.

@@

Recent studies~\cite{chen12b,fang15} have used user inputs to perform {\itercomp} in distributed data centers. Using these approaches, each
compilation worker receives a subset of the input dataset to evaluate a small set of optimization settings on. The best such setting from
each round is used for subsequent executions of the same code. The best-found compilation strategy is then re-fined over time, by testing
new settings and re-evaluating old ones on new datasets. Like in~\cite{chen10} every setting is tested on the same set of inputs, wasting
time and energy. This approach only works well with MapReduce-like workloads, since it relies on the framework for repeating the same
computation multiple times without causing side-effects.

Fursin~\etal~\cite{fursin07} attempted to find a general approach for online \itercomp. They proposed to use instructions per cycle (IPC)
as a metric for comparing different runs of the same program with different optimization settings and different inputs. Their hypothesis is
that application binaries displaying higher IPC are more efficient than lower IPC binaries. However, many optimization techniques affect
IPC in the opposite way, making the program more efficient while lowering IPC. We have also shown in Section~\ref{sec:motivation}, IPC is
not strongly correlate to speedup. Furthermore, IPC cannot be used to estimate performance for multi-threaded
applications~\cite{alameldeen06,eyerman08} due to the side-effect of thread contention. Aalameldeen and Wood~\cite{alameldeen06} also
suggest that relying on IPC can be harmful, and work metrics like time per work unit or work units per time should be used. However, prior
works all rely on developers to provide such metrics. This paper presents the first automatic approach to derive a such work metric from
arbitrary code without developer intervention.
